---
title: "Task_07_notebook"
author: "Rose Borden"
date: "2/22/2017"
output: pdf_document
---
This code sets some options for how the chunks will be displayed and saved. Echo =
TRUE means the R code will be saved when the document is knitted. fig.path tells it
where to save the figures to.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.path = "images/")
```

This code tells R to load the nycflights13 dataset, the babynames dataset, and the
tidyverse packages.
```{r}
library(nycflights13)
library(tidyverse)
library(babynames)
```

This code makes a table called "wind" by selecting only the origin, wind direction,
and wind speed columns from the weather dataset in nycflights13.
```{r}
wind <- nycflights13::weather %>%
  select(origin, wind_dir, wind_speed)
print(wind)
```

This code uses ggplot2 to make a plot of wind speed vs the airport of origin to determine if there are any clear outliers that should be removed. There is one point
over 1000 mph which should be removed before further operations.
```{r}
ggplot(wind) + geom_point(aes(x = origin, y = wind_speed)) 
```

This code filters the wind speed data to remove the point over 1000 mph. 
```{r}
wind1 <- wind %>%
  filter(wind_speed<1000)
head(wind1)
```

This code makes a plot of wind speed vs origin with the filtered data to see if there are any more outliers. There are not.
```{r}
ggplot(wind1) + geom_point(aes(x = origin, y = wind_speed))
```

This code makes a new table of the median wind speed for each direction at each
airport. First, we group the data first by airport of origin and and then by wind
direction. Then we can calculate the median wind speed, and it will be displayed as
one median speed per wind direction per airport.
```{r}
median_wind <- wind1 %>%
  group_by(origin, wind_dir) %>%
  summarise(med_speed = median(wind_speed, na.rm = TRUE))
median_wind
```

This code plots the median wind speed vs the wind direction, with one plot for each 
airport of origin. 
```{r}
med_wind_plot <- ggplot(median_wind) + geom_point(aes(x = wind_dir, y = med_speed)) + facet_wrap(~origin) + xlab("Wind Direction") + ylab("Median Speed")
print(med_wind_plot)
```

This code makes a table by joining the airlines dataset to the flights dataset using
their common column "carrier". The data are then filtered to only those with the
origin airport of JFK. They are arranged by descending distance, and then selected
so only the airline name and distance columns are retained. 
```{r}
JFK_dist <- left_join(nycflights13::airlines, nycflights13::flights, by = "carrier") %>%
  filter(origin == "JFK") %>%
  arrange(desc(distance)) %>%
  select(name, distance)
```

This code makes a new table by joining the airlines dataset to the flights dataset. 
The data are then filtered to only those points with EWR as the airport of origin. 
They are grouped by airline name and month. The summarize function makes a new 
column named "n" that lists the number of flights that leave EWR each month. The 
spread function then makes the table into a wide format table. 
```{r}
EWR_flights <- left_join(nycflights13::airlines, nycflights13::flights, by = "carrier") %>%
  filter(origin == "EWR") %>%
  group_by(name, month) %>%
  summarize(n=n()) %>%
  spread(key = month, n)
```

This code makes a table using the babynames dataset; it is filtered to only include 
names from the year 2014 and grouped by sex, so that when the top_n function is 
applied it will save only the top 10 names for each sex. 
```{r}
name_freq_14 <- babynames %>%
  filter(year == "2014") %>%
  group_by(sex) %>%
  top_n(10, n)
```


```{r}
name_freq_all <- inner_join(babynames, name_freq_14, by = "name", na.rm = TRUE) %>%
  filter(sex.x == sex.y) %>%
  select(name, year.x, prop.x, sex.y)
```


```{r}
name_plot <- ggplot(name_freq_all, aes(x = year.x, y = prop.x, color = sex.y)) + 
  geom_point() +
  scale_color_manual(values = c("seagreen3", "lightgoldenrod2")) + 
  facet_wrap(~ sex.y + name) + 
  xlab("Year") + 
  ylab("Frequency")
print(name_plot)
```


```{r}
girl_names <- babynames %>%
  filter(sex == "F", year == 1896|year == 1942|year == 2016) %>%
  group_by(year) %>%
  mutate(ranking = dense_rank(desc(n))) %>%
  filter(ranking > 25 & ranking < 30)
print(girl_names)
```

I want to find the number of planes manufactured by certain manufacturers for each year. First, I group the data by the name of the manufacturer and then by the year each plane was made. I then use the summarise function to make a column counting the number of planes per year per manufacturer. Then I arrange them in ascending order by year so it is easy to read.
```{r}
planes_make <- nycflights13::planes %>%
  group_by(manufacturer, year) %>%
  summarise(n = n()) %>%
  select(manufacturer, year, n) %>%
  arrange(year)
print(planes_make)
```

This code filters the table made above to include only planes manufactured by Boeing, Airbus, and McDonnell Douglas.
```{r}
some_planes <- planes_make %>%
  filter(manufacturer == "BOEING"|manufacturer == "AIRBUS INDUSTRIE"|manufacturer == "MCDONNELL DOUGLAS")
some_planes
```

This code makes a plot of the table in the previous step, showing the number of planes manufactured per year, color coded by manufacturer.
```{r}
planes_plot <- ggplot(some_planes, aes(x = year, y = n, color = manufacturer)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
print(planes_plot)
```

